---
title: "The 2017 Rice Subsidy Policy"
author: "JANNISAH PALO"
format: pdf
editor: visual
warning: FALSE
---

## BEFORE DiD

### Dataset

```{r}
library(dplyr)
library(readr)

# -------------------------
# LOAD FILES
# -------------------------

hh14 <- read_csv("APIS PUF 2014 Household.CSV")
person14 <- read_csv("APIS PUF 2014 Person.CSV")

hh16 <- read_csv("APIS PUF 2016 Household.CSV")
person16 <- read_csv("APIS PUF 2016 Person.CSV")

hh17 <- read_csv("APIS PUF 2017 Household.CSV")
person17 <- read_csv("APIS PUF 2017 Person.CSV")
```

---- 2014 ----

```{r}
person14_hh <- person14 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hh_2014 <- hh14 %>%
  left_join(person14_hh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps)
  ) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  filter(REGN == 10) 
```

---- 2016 ----

```{r}
person16_hh <- person16 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hh_2016 <- hh16 %>%
  left_join(person16_hh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 0
  ) %>%
  filter(REGN == 10)
```

---- 2017 ----

```{r}
person17_hh <- person17 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hh_2017 <- hh17 %>%
  left_join(person17_hh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 1
  ) %>%
  rename(REGN = REGO) %>%
  filter(REGN == 10)
```

### Stack

---- STACK ----

```{r}
hh_2014$year <- 2014
hh_2016$year <- 2016
hh_2017$year <- 2017

hh_2014 <- hh_2014 %>% mutate(across(everything(), as.character))
hh_2016 <- hh_2016 %>% mutate(across(everything(), as.character))
hh_2017 <- hh_2017 %>% mutate(across(everything(), as.character))

combined <- bind_rows(hh_2014, hh_2016, hh_2017)
```

```{r}
combined <- subset(combined, REGN == 10)

library(dplyr)

combined <- combined %>%
  filter(REGN == 10)

with(combined, table(fourps, post))
```

### Prepare variables

```{r}
combined <- combined %>%
  mutate(
    fourps = as.numeric(fourps),
    post = as.numeric(post),
    FOOD = as.numeric(FOOD),
    year = as.numeric(year),
    REGN = as.factor(REGN),
    RFACT = as.numeric(RFACT),
    FSIZE = as.numeric(FSIZE),
    PUFEQ5 = as.numeric(PUFEQ5),
    PUFEQ6G = as.numeric(PUFEQ6G),
    TOINC = as.numeric(TOINC),
    PUFH04_SEX = as.numeric(PUFH04_SEX),
    PUFH05_AGE = as.numeric(PUFH05_AGE)
  )
```

### Use only 2016 and 2017 for the DiD model

```{r}
did_data <- combined %>%
  filter(year %in% c(2016, 2017)) %>%
  filter(!is.na(FOOD), !is.na(fourps), !is.na(post), !is.na(year), !is.na(REGN), !is.na(RFACT), !is.na(FSIZE), !is.na(PUFEQ5), !is.na(PUFEQ6G), !is.na(TOINC), !is.na(PUFH04_SEX), !is.na(PUFH05_AGE))

with(did_data, table(fourps, post))
```

## DiD MODELS

### Basic DiD

```{r}
did_model <- lm(FOOD ~ fourps * post, data = did_data)
summary(did_model)
```

### DiD with covariates & survey weights (Final Model)

```{r}
did_model_w <- lm(
  FOOD ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G ,
  data = did_data,
  weights = RFACT
)
summary(did_model_w)
```

## DiD GRAPH

### Weighted group means

```{r}
trend_dataa <- combined %>%
  filter(year %in% c(2016, 2017)) %>%
  filter(!is.na(FOOD), !is.na(fourps), !is.na(RFACT)) %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, w = RFACT, na.rm = TRUE),
    .groups = "drop"
  )

trend_dataa$mean_food
```

```{r}
library(dplyr)

# Extract values
treat_2016 <- trend_dataa %>%
  filter(year == 2016, fourps == 1) %>%
  pull(mean_food)

control_2016 <- trend_dataa %>%
  filter(year == 2016, fourps == 0) %>%
  pull(mean_food)

control_2017 <- trend_dataa %>%
  filter(year == 2017, fourps == 0) %>%
  pull(mean_food)

# Control group change
control_change <- control_2017 - control_2016

# Counterfactual for treated
cf_2017 <- treat_2016 + control_change

counterfactual <- data.frame(
  year = c(2016, 2017),
  mean_food = c(treat_2016, cf_2017)
)
```

### Plot

```{r}
treat_2017 <- trend_dataa %>%
  filter(year == 2017, fourps == 1) %>%
  pull(mean_food)

trend_dataa$mean_food
```

```{r}
library(ggplot2)

ggplot(trend_dataa,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +

  geom_line() +
  geom_point() +

  geom_line(data = counterfactual,
            aes(x = year, y = mean_food),
            linetype = "dashed",
            color = "black",
            inherit.aes = FALSE) +
  
  geom_segment(
    aes(x = 2017, xend = 2017,
        y = cf_2017, yend = treat_2017),
    color = "#FF69B4",
    inherit.aes = FALSE
  ) +
  
  scale_x_continuous(breaks = c(2016, 2017)) +

  labs(
    x = "Year",
    y = "Weighted Mean Food Expenditure",
    color = "Group",
    title = "Food Expenditure Trends of 4Ps and Non-4Ps Households"
  ) +
  scale_color_discrete(labels = c("Non-4Ps", "4Ps")) +
theme_minimal() +
  
 
  theme(
    plot.title = element_text(size = 16, face = "bold"), 
    axis.title = element_text(size = 14),                
    axis.text = element_text(size = 12),                 
    legend.title = element_text(size = 13),              
    legend.text = element_text(size = 12)                
  )
```

```{r}
ggplot(trend_dataa,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +
  
  # Main lines - thicker
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5, shape = 21, 
             fill = "white", stroke = 1.5) +
  
  # Counterfactual dashed line
  geom_line(data = counterfactual,
            aes(x = year, y = mean_food),
            linetype = "dashed",
            color = "black",
            linewidth = 1,
            inherit.aes = FALSE) +
  
  # DiD gap segment
  geom_segment(
    aes(x = 2017, xend = 2017,
        y = cf_2017, yend = treat_2017),
    color = "black",
    linewidth = 0.8,
    linetype = "solid",
    arrow = arrow(ends = "both",        
                  length = unit(0.2, "cm")),
    inherit.aes = FALSE
  ) +
  
  # DiD label on the gap
  annotate("text",
           x = 2017.05,
           y = (cf_2017 + treat_2017) / 2,
           label = "DiD = ₱2,753",
           hjust = 0,
           size = 4.5,            
           color = "black") +
  
  # Counterfactual label
  annotate("text",
           x = 2016.5,
           y = cf_2017 - 500,
           label = "Counterfactual",
           hjust = 0,
           size = 4,              
           color = "black",
           fontface = "italic") +
  
  # Colors
  scale_color_manual(
    values = c("0" = "#E07B6A",    
               "1" = "#5B9BD5"),   
    labels = c("Non-4Ps", "4Ps")
  ) +
  
  # Axis
  scale_x_continuous(
    breaks = c(2016, 2017),
    labels = c("2016\n(Pre)", "2017\n(Post)")
  ) +
  
  scale_y_continuous(
    labels = scales::comma_format(
      prefix = "₱"
    )
  ) +
  
  # Labels
  labs(
    x = "Survey Year",
    y = "Weighted Mean Food Expenditure (₱)",
    color = "Household Group",
    title = "Difference-in-Differences: Food Expenditure Trends",
    subtitle = "4Ps and Non-4Ps Households, Region X (2016–2017)",
    caption = "Note: Dashed line represents counterfactual trend for 4Ps households.\nSource: APIS 2016 and 2017, Philippine Statistics Authority."
  ) +
  
  # Theme
  theme_classic() +
  theme(
    # Title & Captions
    plot.title = element_text(face = "bold", 
                              size = 16,          
                              hjust = 0),
    plot.subtitle = element_text(size = 12,       
                                 color = "gray40",
                                 hjust = 0),
    plot.caption = element_text(size = 9,        
                                color = "gray50",
                                hjust = 0),
    
    # Axes
    axis.title = element_text(size = 13,          
                              face = "bold"),
    axis.text = element_text(size = 11),          
    axis.line = element_line(color = "black",
                             linewidth = 0.5),
    
    # Legend
    legend.position = "right",
    legend.title = element_text(face = "bold",
                                size = 12),       
    legend.text = element_text(size = 11),       
    legend.background = element_rect(
      fill = "white", color = "gray80"
    ),
    
    # Grid
    panel.grid.major.y = element_line(
      color = "gray90", linewidth = 0.5
    ),
    panel.grid.major.x = element_blank(),
    
    # Margins
    plot.margin = margin(15, 15, 15, 15)
  )

# Save high resolution for thesis
ggsave("did_plot.png", 
       width = 7, height = 5, 
       dpi = 300,
       bg = "white")
```

## DiD GRAPH

```{r}
trend_data <- combined %>%
  filter(year %in% c(2014, 2016, 2017)) %>%
  filter(!is.na(FOOD), !is.na(fourps), !is.na(RFACT)) %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, w = RFACT, na.rm = TRUE),
    .groups = "drop"
  )

trend_data$mean_food
```

```{r}
library(dplyr)

# Extract values
treat_2016 <- trend_data %>%
  filter(year == 2016, fourps == 1) %>%
  pull(mean_food)

control_2016 <- trend_data %>%
  filter(year == 2016, fourps == 0) %>%
  pull(mean_food)

control_2017 <- trend_data %>%
  filter(year == 2017, fourps == 0) %>%
  pull(mean_food)

# Control group change
control_change <- control_2017 - control_2016

# Counterfactual for treated
cf_2017 <- treat_2016 + control_change

counterfactual <- data.frame(
  year = c(2016, 2017),
  mean_food = c(treat_2016, cf_2017)
)
```

```{r}
ggplot(trend_data,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +

  geom_line() +
  geom_point() +

  geom_line(data = counterfactual,
            aes(x = year, y = mean_food),
            linetype = "dashed",
            color = "black",
            inherit.aes = FALSE) +
  
  geom_segment(
    aes(x = 2017, xend = 2017,
        y = cf_2017, yend = treat_2017),
    color = "#FF69B4",
    inherit.aes = FALSE
  ) +

  labs(
    x = "Year",
    y = "Weighted Mean Food Expenditure",
    color = "Group",
    title = "Food Expenditure Trends of 4Ps and Non-4Ps Households"
  ) +
  scale_color_discrete(labels = c("Non-4Ps", "4Ps")) +
  
  theme_minimal() +
  
 
  theme(
    plot.title = element_text(size = 16, face = "bold"), 
    axis.title = element_text(size = 14, face = "bold"), 
    axis.text = element_text(size = 12),                 
    legend.title = element_text(size = 13, face = "bold"),
    legend.text = element_text(size = 12)                
  )
```

```{r}
hh11 <- suppressMessages(read_csv("APIS PUF 2011 Household.CSV"))
person11 <- read_csv("APIS PUF 2011 Person.CSV")

#2011
person11_hhh <- person11 %>%
  rename(ID = PHHNUM) %>%
  rename(PPS01 = PPS04) %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2011 <- hh11 %>%
  rename(ID = PHHNUM) %>%
  left_join(person11_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2011
  ) %>%
  rename(REGN = PREG) %>%
  rename(FSIZE = PUFSIZE) %>%
  rename(RFACT = HHWGT) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(FOOD = PUFFOOD) %>%
  rename(TOINC = PUFTOTINC) %>%
  rename(PUFH12_HGC = PUFH14_HGC) %>%
  filter(REGN == 10)

hh13 <- read_csv("APIS PUF 2013 Household.CSV")
person13 <- read_csv("APIS PUF 2013 Person.CSV")

#2013
person13_hhh <- person13 %>%
  rename(ID = SEQ) %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2013 <- hh13 %>%
  rename(ID = SEQ) %>%
  left_join(person13_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2013
  ) %>%
  rename(REGN = REGION) %>%
  rename(FOOD = T_FOOD) %>%
  rename(PUFH12_HGC = PUFH14_HGC) %>%
  filter(REGN == 10)

#2014
person14_hhh <- person14 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2014 <- hh14 %>%
  left_join(person14_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2014
  ) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  filter(REGN == 10) 

#2016
person16_hhh <- person16 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2016 <- hh16 %>%
  left_join(person16_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 0,
    year = 2016
  ) %>%
  filter(REGN == 10)

#2017
person17_hhh <- person17 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2017 <- hh17 %>%
  left_join(person17_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 1,
    year = 2017
  ) %>%
  rename(REGN = REGO) %>%
  filter(REGN == 10)

hhh_2011$year <- 2011
hhh_2013$year <- 2013
hhh_2014$year <- 2014
hhh_2016$year <- 2016
hhh_2017$year <- 2017

hhh_2011 <- hhh_2011 %>% mutate(across(everything(), as.character))
hhh_2013 <- hhh_2013 %>% mutate(across(everything(), as.character))
hhh_2014 <- hhh_2014 %>% mutate(across(everything(), as.character))
hhh_2016 <- hhh_2016 %>% mutate(across(everything(), as.character))
hhh_2017 <- hhh_2017 %>% mutate(across(everything(), as.character))

combined5 <- bind_rows(hhh_2011, hhh_2013, hhh_2014, hhh_2016, hhh_2017)

combined5 <- subset(combined5, REGN == 10)
library(dplyr)
combined5 <- combined5 %>%
  filter(REGN == 10)
with(combined5, table(fourps, post))

combined5 <- combined5 %>%
  mutate(
    fourps = as.numeric(fourps),
    post = as.numeric(post),
    FOOD = as.numeric(FOOD),
    year = as.numeric(year),
    REGN = as.factor(REGN),
    RFACT = as.numeric(RFACT),
    FSIZE = as.numeric(FSIZE),
    PUFEQ5 = as.numeric(PUFEQ5),
    PUFEQ6G = as.numeric(PUFEQ6G),
    TOINC = as.numeric(TOINC),
    PUFH04_SEX = as.numeric(PUFH04_SEX),
    PUFH05_AGE = as.numeric(PUFH05_AGE)
  )
```

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(scales)

# First make sure trend_data includes all years
trend_data <- combined5 %>%
  filter(year %in% c(2011, 2013, 2014, 2016, 2017)) %>%
  filter(!is.na(FOOD), !is.na(fourps), !is.na(RFACT)) %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, w = RFACT, na.rm = TRUE),
    .groups = "drop"
  )

# Counterfactual
treat_2016  <- trend_data %>% 
  filter(year == 2016, fourps == 1) %>% pull(mean_food)
control_2016 <- trend_data %>% 
  filter(year == 2016, fourps == 0) %>% pull(mean_food)
control_2017 <- trend_data %>% 
  filter(year == 2017, fourps == 0) %>% pull(mean_food)
treat_2017  <- trend_data %>% 
  filter(year == 2017, fourps == 1) %>% pull(mean_food)

control_change <- control_2017 - control_2016
cf_2017 <- treat_2016 + control_change

counterfactual <- data.frame(
  year     = c(2016, 2017),
  mean_food = c(treat_2016, cf_2017)
)

# Plot
ggplot(trend_data,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +
  
  # Vertical intervention line
  geom_vline(xintercept = 2016.5,
             linetype = "dotted",
             color = "gray40",
             linewidth = 0.8) +
  
  # Pre/Post shading
  annotate("rect",
           xmin = 2010.5, xmax = 2016.5,
           ymin = -Inf, ymax = Inf,
           fill = "gray95", alpha = 0.5) +
  annotate("rect",
           xmin = 2016.5, xmax = 2017.5,
           ymin = -Inf, ymax = Inf,
           fill = "lightyellow", alpha = 0.5) +
  
  # Pre/Post labels
  annotate("text", x = 2013.5, y = 38500,
           label = "Pre-Intervention",
           size = 3.2, color = "gray40",
           fontface = "italic") +
  annotate("text", x = 2017, y = 38500,
           label = "Post",
           size = 3.2, color = "gray40",
           fontface = "italic") +
  
  # Main lines
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5, shape = 21,
             fill = "white", stroke = 1.5) +
  
  # Counterfactual line (2016 to 2017 only)
  geom_line(data = counterfactual,
            aes(x = year, y = mean_food),
            linetype = "dashed",
            color = "black",
            linewidth = 1,
            inherit.aes = FALSE) +
  
  # DiD gap arrow
  geom_segment(
    aes(x = 2017, xend = 2017,
        y = cf_2017, yend = treat_2017),
    color = "black",
    linewidth = 0.8,
    arrow = arrow(ends = "both",
                  length = unit(0.2, "cm")),
    inherit.aes = FALSE
  ) +
  
  # DiD estimate label
  annotate("text",
           x = 2017.05,
           y = (cf_2017 + treat_2017) / 2,
           label = paste0("DiD = ₱", 
                         format(round(treat_2017 - cf_2017, 0),
                                big.mark = ",")),
           hjust = 0,
           size = 3.2,
           color = "black",
           fontface = "bold") +
  
  # Counterfactual label
  annotate("text",
           x = 2016.55,
           y = cf_2017 - 800,
           label = "Counterfactual",
           hjust = 0,
           size = 3,
           color = "black",
           fontface = "italic") +
  
  # Colors
  scale_color_manual(
    values = c("0" = "#E07B6A",
               "1" = "#5B9BD5"),
    labels = c("Non-4Ps", "4Ps")
  ) +
  
  # Axis formatting
  scale_x_continuous(
    breaks = c(2011, 2013, 2014, 2016, 2017)
  ) +
  scale_y_continuous(
    labels = comma_format(prefix = "₱"),
    limits = c(25000, 40000)
  ) +
  
  # Labels
  labs(
    x = "Survey Year",
    y = "Weighted Mean Food Expenditure (₱)",
    color = "Household Group",
    title = "Food Expenditure Trends of 4Ps and Non-4Ps Households",
    subtitle = "Region X, APIS 2011–2017 | Dashed line = counterfactual trend",
    caption = "Source: APIS 2011–2017, Philippine Statistics Authority.\nNote: Weighted means using RFACT survey expansion factors."
  ) +
  
  # Theme
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold",
                              size = 12,
                              hjust = 0),
    plot.subtitle = element_text(size = 9,
                                 color = "gray40",
                                 hjust = 0),
    plot.caption = element_text(size = 8,
                                color = "gray50",
                                hjust = 0),
    axis.title = element_text(size = 11,
                              face = "bold"),
    axis.text = element_text(size = 10),
    axis.line = element_line(color = "black",
                             linewidth = 0.5),
    legend.position = "right",
    legend.title = element_text(face = "bold",
                                size = 10),
    legend.text = element_text(size = 10),
    legend.background = element_rect(
      fill = "white", color = "gray80"
    ),
    panel.grid.major.y = element_line(
      color = "gray90", linewidth = 0.4
    ),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 20, 15, 15)
  )

# Save
ggsave("did_pretrends_plot.png",
       width = 8, height = 5,
       dpi = 300,
       bg = "white")
```

## FINAL Exploratory Data Analysis + DiD + Robustness Check

#Descriptive Statistics

## PART 1: Data Description

```{r}
# Sample description
table(combined5$year)

# 2×2 DiD structure (2016–2017)
did_sample <- combined5 %>% filter(year %in% c(2016, 2017))
table(did_sample$fourps, did_sample$post)
```

## EDA

## PART 2: Distributional Properties

```{r}
library(ggplot2)

#manually construct, what is the range?????

# FOOD distribution
ggplot(combined5, aes(x = FOOD)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Food Expenditure")

# TOINC
ggplot(combined5, aes(x = TOINC)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Income")
```

```{r, echo=FALSE}
library(ggplot2)
library(scales)

# FOOD histogram
ggplot(combined5, aes(x = FOOD)) +
  geom_histogram(bins = 50,
                 fill = "#5B9BD5",
                 color = "white",
                 linewidth = 0.3) +
  
  # Mean and median lines
  geom_vline(aes(xintercept = mean(FOOD, na.rm = TRUE)),
             color = "#E07B6A",
             linetype = "dashed",
             linewidth = 0.9) +
  geom_vline(aes(xintercept = median(FOOD, na.rm = TRUE)),
             color = "darkgreen",
             linetype = "dashed",
             linewidth = 0.9) +
  
  # Mean and median labels
  annotate("text",
           x = mean(combined5$FOOD, na.rm = TRUE) + 8000,
           y = 900,
           label = paste0("Mean = ₱",
                         format(round(mean(combined5$FOOD,
                                         na.rm = TRUE), 0),
                                big.mark = ",")),
           color = "#E07B6A",
           size = 3.2,
           hjust = 0) +
  annotate("text",
           x = median(combined5$FOOD, na.rm = TRUE) + 8000,
           y = 820,
           label = paste0("Median = ₱",
                         format(round(median(combined5$FOOD,
                                           na.rm = TRUE), 0),
                                big.mark = ",")),
           color = "darkgreen",
           size = 3.2,
           hjust = 0) +
  
  # Axis formatting
  scale_x_continuous(
    labels = comma_format(prefix = "₱"),
    breaks = seq(0, 500000, by = 100000)
  ) +
  scale_y_continuous(
    labels = comma_format()
  ) +
  
  # Labels
  labs(
    title = "Distribution of Household Food Expenditure",
    subtitle = "Region X, APIS 2011–2017 (n = 3,995)",
    x = "Annual Food Expenditure (₱)",
    y = "Number of Households",
    caption = "Note: Distribution is right-skewed with extreme values up to ₱476,060.\nSource: APIS 2011–2017, Philippine Statistics Authority."
  ) +
  
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold",
                              size = 12,
                              hjust = 0),
    plot.subtitle = element_text(size = 9,
                                 color = "gray40",
                                 hjust = 0),
    plot.caption = element_text(size = 8,
                                color = "gray50",
                                hjust = 0),
    axis.title = element_text(size = 11,
                              face = "bold"),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 30,
                               hjust = 1),
    panel.grid.major.y = element_line(color = "gray90",
                                      linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 15, 15, 15)
  )

ggsave("food_distribution.png",
       width = 7, height = 5,
       dpi = 300, bg = "white")


# TOINC histogram
ggplot(combined5, aes(x = TOINC)) +
  geom_histogram(bins = 50,
                 fill = "#5B9BD5",
                 color = "white",
                 linewidth = 0.3) +
  
  # Mean and median lines
  geom_vline(aes(xintercept = mean(TOINC, na.rm = TRUE)),
             color = "#E07B6A",
             linetype = "dashed",
             linewidth = 0.9) +
  geom_vline(aes(xintercept = median(TOINC, na.rm = TRUE)),
             color = "darkgreen",
             linetype = "dashed",
             linewidth = 0.9) +
  
  # Mean and median labels
  annotate("text",
           x = mean(combined5$TOINC, na.rm = TRUE) + 50000,
           y = 1500,
           label = paste0("Mean = ₱",
                         format(round(mean(combined5$TOINC,
                                         na.rm = TRUE), 0),
                                big.mark = ",")),
           color = "#E07B6A",
           size = 3.2,
           hjust = 0) +
  annotate("text",
           x = median(combined5$TOINC, na.rm = TRUE) + 50000,
           y = 1380,
           label = paste0("Median = ₱",
                         format(round(median(combined5$TOINC,
                                           na.rm = TRUE), 0),
                                big.mark = ",")),
           color = "darkgreen",
           size = 3.2,
           hjust = 0) +
  
  # Axis formatting
  scale_x_continuous(
    labels = comma_format(prefix = "₱"),
    breaks = seq(0, 5000000, by = 1000000)
  ) +
  scale_y_continuous(
    labels = comma_format()
  ) +
  
  # Labels
  labs(
    title = "Distribution of Household Total Income",
    subtitle = "Region X, APIS 2011–2017 (n = 3,995)",
    x = "Annual Total Income (₱)",
    y = "Number of Households",
    caption = "Note: Distribution is heavily right-skewed, justifying log transformation in regression.\nSource: APIS 2011–2017, Philippine Statistics Authority."
  ) +
  
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold",
                              size = 12,
                              hjust = 0),
    plot.subtitle = element_text(size = 9,
                                 color = "gray40",
                                 hjust = 0),
    plot.caption = element_text(size = 8,
                                color = "gray50",
                                hjust = 0),
    axis.title = element_text(size = 11,
                              face = "bold"),
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 30,
                               hjust = 1),
    panel.grid.major.y = element_line(color = "gray90",
                                      linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 15, 15, 15)
  )

ggsave("income_distribution.png",
       width = 7, height = 5,
       dpi = 300, bg = "white")

#check if 5,000,000 is present in csv
```

## PART 3: Group Differences

```{r}
#OUTLIER DETECTION: BOXPLOT

# Boxplot of FOOD by 4Ps status
#ggplot(combined5, aes(x = factor(fourps), y = FOOD, color = factor(fourps))) +
#  geom_boxplot() +
#  labs(x = "4Ps Status", y = "Food Expenditure", title = "Distribution of Total Food Expenditure by 4Ps Status")

# Boxplot of TOINC by 4Ps status
#ggplot(combined5, aes(x = factor(fourps), y = TOINC, color = factor(fourps))) +
#  geom_boxplot() +
#  labs(x = "4Ps Status", y = "Total Income",
#       title = "Distribution of Total Income by 4Ps Status")



# Tells that both groups are statistically different and validates for controlling for them as covariates

# Balance test (2016)
pre_data <- combined5 %>% filter(year == 2016)
t.test(FSIZE ~ fourps, data = pre_data)
t.test(TOINC ~ fourps, data = pre_data)
t.test(PUFEQ6G ~ fourps, data = pre_data)
```

## PART 4: Relationships Among Variables

```{r}
# Correlation
combined5 %>%
  select(FOOD, TOINC, FSIZE, PUFEQ6G) %>%
  cor(use = "complete.obs")
```

## PART 5: DiD Assumption Checks

## Parallel Trend Assumption

### PRE-TREND ANALYSIS

Graph:

```{r}
# Parallel trends plot
trend_data <- combined5 %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, RFACT, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(trend_data,
       aes(x = year, y = mean_food,
           color = factor(fourps), group = fourps)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Year",
    y = "Weighted Mean Food Expenditure",
    color = "Group",
    title = "Food Expenditure Trends of 4Ps and Non-4Ps Households"
  ) +
  scale_color_discrete(labels = c("Non-4Ps", "4Ps")) +
  theme_minimal() +
  
  
  theme(
    plot.title = element_text(size = 16, face = "bold"),   
    axis.title = element_text(size = 14, face = "bold"),   
    axis.text = element_text(size = 12),                   
    legend.title = element_text(size = 13, face = "bold"), 
    legend.text = element_text(size = 12)                  
  )
```

```{r, echo=FALSE}
library(ggplot2)
library(scales)
library(dplyr)

# Data
trend_data <- combined5 %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, RFACT, na.rm = TRUE),
    .groups = "drop"
  )

# Plot
ggplot(trend_data,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +
  
  # ── Shading ──────────────────────────────────────
  annotate("rect",
           xmin = 2010.5, xmax = 2017.5,
           ymin = -Inf,   ymax = Inf,
           fill = "gray95", alpha = 0.6) +
#  annotate("rect",
#           xmin = 2016.5, xmax = 2017.5,
#           ymin = -Inf,   ymax = Inf,
#           fill = "lightyellow", alpha = 0.6) +

  # ── Intervention line ─────────────────────────────
#  geom_vline(xintercept = 2016.5,
#             linetype = "dotted",
#             color = "gray40",
#             linewidth = 0.8) +

  # ── Period labels ─────────────────────────────────
  annotate("text",
           x = 2013, y = 38500,
           label = "Pre-Intervention",
           size = 3.2, color = "gray40",
           fontface = "italic") +
  annotate("text",
           x = 2017, y = 38500,
           label = "Post",
           size = 3.2, color = "gray40",
           fontface = "italic") +

  # ── Lines and points ──────────────────────────────
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5,
             shape = 21,
             fill = "white",
             stroke = 1.5) +

  # ── Value labels on each point ────────────────────
  geom_text(aes(label = paste0("₱",
                               format(round(mean_food, 0),
                                      big.mark = ","))),
            vjust = -1.2,
            size = 2.8,
            show.legend = FALSE) +

  # ── Colors ────────────────────────────────────────
  scale_color_manual(
    values = c("0" = "#E07B6A",
               "1" = "#5B9BD5"),
    labels = c("Non-4Ps (Control)",
               "4Ps (Treated)")
  ) +

  # ── Axes ──────────────────────────────────────────
  scale_x_continuous(
    breaks = c(2011, 2013, 2014, 2016, 2017)
  ) +
  scale_y_continuous(
    labels = comma_format(prefix = "₱"),
    limits = c(25000, 40500),
    breaks = seq(25000, 40000, by = 2500)
  ) +

  # ── Labels ────────────────────────────────────────
  labs(
    title    = "Parallel Trends in Food Expenditure: 4Ps vs Non-4Ps Households",
    subtitle = "Region X, APIS 2011–2017 | Weighted by RFACT survey expansion factors",
    x        = "Survey Year",
    y        = "Weighted Mean Food Expenditure (₱)",
    color    = "Household Group",
    caption  = paste0(
      "Note: Both groups exhibit broadly similar pre-intervention trends (2011–2016),\n",
      "supporting the parallel trends assumption underlying the DiD estimator.\n",
      "Source: APIS 2011–2017, Philippine Statistics Authority."
    )
  ) +

  # ── Theme ─────────────────────────────────────────
  theme_classic() +
  theme(
    plot.title    = element_text(face = "bold",
                                 size = 12,
                                 hjust = 0),
    plot.subtitle = element_text(size = 9,
                                 color = "gray40",
                                 hjust = 0),
    plot.caption  = element_text(size = 8,
                                 color = "gray50",
                                 hjust = 0),
    axis.title    = element_text(size = 11,
                                 face = "bold"),
    axis.text     = element_text(size = 10),
    legend.position    = "right",
    legend.title       = element_text(face = "bold",
                                      size = 10),
    legend.text        = element_text(size = 10),
    legend.background  = element_rect(fill = "white",
                                      color = "gray80"),
    panel.grid.major.y = element_line(color = "gray90",
                                      linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(15, 20, 15, 15)
  )

ggsave("parallel_trends_full.png",
       width = 8, height = 5,
       dpi = 300, bg = "white")
```

```{r, echo=FALSE}

library(ggplot2)
library(scales)
library(dplyr)
library(ggrepel)

# Data
trend_data <- combined5 %>%
  group_by(year, fourps) %>%
  summarise(
    mean_food = weighted.mean(FOOD, RFACT, na.rm = TRUE),
    .groups = "drop"
  )

# Plot
ggplot(trend_data,
       aes(x = year, y = mean_food,
           color = factor(fourps),
           group = factor(fourps))) +
  
  # ── Shading ──────────────────────────────────────
  annotate("rect",
           xmin = 2010.5, xmax = 2017.5,
           ymin = -Inf,   ymax = Inf,
           fill = "gray95", alpha = 0.6) +

  # ── Lines and points ──────────────────────────────
  geom_line(linewidth = 1.2) +
  geom_point(size = 3.5,
             shape = 21,
             fill = "white",
             stroke = 1.5) +

  # ── Clean, Non-overlapping Value Labels ───────────
  # Force labels to stack directly above/below the points
  geom_text_repel(aes(label = paste0("₱", 
                               format(round(mean_food, 0),
                                      big.mark = ","))),
                  size = 3.5,
                  show.legend = FALSE,
                  direction = "y",         
                  nudge_y = 400,            
                  box.padding = 0.5,       
                  point.padding = 0.3,     
                  force = 2) + 

  # ── Colors ────────────────────────────────────────
  scale_color_manual(
    values = c("0" = "#E07B6A",
               "1" = "#5B9BD5"),
    labels = c("Non-4Ps (Control)",
               "4Ps (Treated)")
  ) +

  # ── Axes ──────────────────────────────────────────
  scale_x_continuous(
    breaks = c(2011, 2013, 2014, 2016, 2017)
  ) +
  scale_y_continuous(
    labels = comma_format(prefix = "₱"),
    limits = c(25000, 43000),                
    breaks = seq(25000, 40000, by = 2500)
  ) +

  # ── Period Labels Safely Anchored at the Top Edge ──
  
  annotate("text", x = 2013.5, y = Inf, label = "PRE-INTERVENTION", 
           vjust = 2, size = 3.8, color = "gray40", fontface = "italic") +
  annotate("text", x = 2017, y = Inf, label = "POST", 
           vjust = 2, size = 3.8, color = "gray40", fontface = "italic") +

  # ── Labels ────────────────────────────────────────
  labs(
    title    = "Parallel Trends in Food Expenditure: 4Ps vs Non-4Ps Households",
    subtitle = "Region X, APIS 2011–2017 | Weighted by RFACT survey expansion factors",
    x        = "Survey Year",
    y        = "Weighted Mean Food Expenditure (₱)",
    color    = "Household Group",
    caption  = paste0(
      "Note: Both groups exhibit broadly similar pre-intervention trends (2011–2016),\n",
      "supporting the parallel trends assumption underlying the DiD estimator.\n",
      "Source: APIS 2011–2017, Philippine Statistics Authority."
    )
  ) +

  # ── Theme ─────────────────────────────────────────
  theme_classic() +
  theme(
    # Titles & Captions
    plot.title    = element_text(face = "bold", size = 15, hjust = 0),
    plot.subtitle = element_text(size = 11, color = "gray40", hjust = 0, margin = margin(b = 15)),
    plot.caption  = element_text(size = 9, color = "gray50", hjust = 0),
    
    # Axes Labels & Values
    axis.title    = element_text(size = 13, face = "bold"),
    axis.text     = element_text(size = 11),
    
    # Legend Elements
    legend.position    = "right",
    legend.title       = element_text(face = "bold", size = 11),
    legend.text        = element_text(size = 11),
    legend.background  = element_rect(fill = "white", color = "gray80"),
    
    # Gridlines and Background
    panel.grid.major.y = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    plot.margin        = margin(15, 20, 15, 15)
  )

ggsave("parallel_trends_full.png",
       width = 8, height = 5,
       dpi = 300, bg = "white")
```

### PLACEBO TEST (PLACEBO DiD)

```{r}
person14_hhh <- person14 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2014 <- hh14 %>%
  left_join(person14_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 0,
    year = 2014
  ) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  filter(REGN == 10) 

person16_hhh <- person16 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2016 <- hh16 %>%
  left_join(person16_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 1,
    year = 2016
  ) %>%
  filter(REGN == 10)

person17_hhh <- person17 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2017 <- hh17 %>%
  left_join(person17_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2017
  ) %>%
  rename(REGN = REGO) %>%
  filter(REGN == 10)

hhh_2014$year <- 2014
hhh_2016$year <- 2016
hhh_2017$year <- 2017

hhh_2014 <- hhh_2014 %>% mutate(across(everything(), as.character))
hhh_2016 <- hhh_2016 %>% mutate(across(everything(), as.character))
hhh_2017 <- hhh_2017 %>% mutate(across(everything(), as.character))

combinedd <- bind_rows(hhh_2014, hhh_2016, hhh_2017)

combinedd <- subset(combinedd, REGN == 10)
library(dplyr)
combinedd <- combinedd %>%
  filter(REGN == 10)
with(combinedd, table(fourps, post))

combinedd <- combinedd %>%
  mutate(
    fourps = as.numeric(fourps),
    post = as.numeric(post),
    FOOD = as.numeric(FOOD),
    year = as.numeric(year),
    REGN = as.factor(REGN),
    RFACT = as.numeric(RFACT),
    FSIZE = as.numeric(FSIZE),
    PUFEQ5 = as.numeric(PUFEQ5),
    PUFEQ6G = as.numeric(PUFEQ6G),
    TOINC = as.numeric(TOINC),
    PUFH04_SEX = as.numeric(PUFH04_SEX),
    PUFH05_AGE = as.numeric(PUFH05_AGE)
  )

trend_did_data <- combinedd %>%
  filter(year %in% c(2014, 2016)) %>%
  filter(!is.na(FOOD), !is.na(fourps), !is.na(post), !is.na(year), !is.na(REGN), !is.na(RFACT), !is.na(FSIZE), !is.na(PUFEQ5), !is.na(PUFEQ6G), !is.na(TOINC), !is.na(PUFH04_SEX), !is.na(PUFH05_AGE))
```

#### DiD 2014 & 2016

```{r}
did_model_ww <- lm(
  FOOD ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G ,
  data = trend_did_data,
  weights = RFACT
)
summary(did_model_ww)
```

### EVENT-STUDY MODEL (Event-Study Regression)

```{r}
hh11 <- suppressMessages(read_csv("APIS PUF 2011 Household.CSV"))
person11 <- read_csv("APIS PUF 2011 Person.CSV")

#2011
person11_hhh <- person11 %>%
  rename(ID = PHHNUM) %>%
  rename(PPS01 = PPS04) %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2011 <- hh11 %>%
  rename(ID = PHHNUM) %>%
  left_join(person11_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2011
  ) %>%
  rename(REGN = PREG) %>%
  rename(FSIZE = PUFSIZE) %>%
  rename(RFACT = HHWGT) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(FOOD = PUFFOOD) %>%
  rename(TOINC = PUFTOTINC) %>%
  rename(PUFH12_HGC = PUFH14_HGC) %>%
  filter(REGN == 10)

hh13 <- read_csv("APIS PUF 2013 Household.CSV")
person13 <- read_csv("APIS PUF 2013 Person.CSV")

#2013
person13_hhh <- person13 %>%
  rename(ID = SEQ) %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2013 <- hh13 %>%
  rename(ID = SEQ) %>%
  left_join(person13_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2013
  ) %>%
  rename(REGN = REGION) %>%
  rename(FOOD = T_FOOD) %>%
  rename(PUFH12_HGC = PUFH14_HGC) %>%
  filter(REGN == 10)

#2014
person14_hhh <- person14 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2014 <- hh14 %>%
  left_join(person14_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    year = 2014
  ) %>%
  rename(PUFEQ6G = PUFFQ6G) %>%
  rename(PUFEQ5 = PUFFQ5) %>%
  filter(REGN == 10) 

#2016
person16_hhh <- person16 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2016 <- hh16 %>%
  left_join(person16_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 0,
    year = 2016
  ) %>%
  filter(REGN == 10)

#2017
person17_hhh <- person17 %>%
  group_by(ID) %>%
  summarise(fourps = as.integer(any(PPS01 == 1, na.rm = TRUE)), .groups = "drop")

hhh_2017 <- hh17 %>%
  left_join(person17_hhh, by = "ID") %>%
  mutate(
    fourps = ifelse(is.na(fourps), 0, fourps),
    post = 1,
    year = 2017
  ) %>%
  rename(REGN = REGO) %>%
  filter(REGN == 10)

hhh_2011$year <- 2011
hhh_2013$year <- 2013
hhh_2014$year <- 2014
hhh_2016$year <- 2016
hhh_2017$year <- 2017

hhh_2011 <- hhh_2011 %>% mutate(across(everything(), as.character))
hhh_2013 <- hhh_2013 %>% mutate(across(everything(), as.character))
hhh_2014 <- hhh_2014 %>% mutate(across(everything(), as.character))
hhh_2016 <- hhh_2016 %>% mutate(across(everything(), as.character))
hhh_2017 <- hhh_2017 %>% mutate(across(everything(), as.character))

combined5 <- bind_rows(hhh_2011, hhh_2013, hhh_2014, hhh_2016, hhh_2017)

combined5 <- subset(combined5, REGN == 10)
library(dplyr)
combined5 <- combined5 %>%
  filter(REGN == 10)
with(combined5, table(fourps, post))

combined5 <- combined5 %>%
  mutate(
    fourps = as.numeric(fourps),
    post = as.numeric(post),
    FOOD = as.numeric(FOOD),
    year = as.numeric(year),
    REGN = as.factor(REGN),
    RFACT = as.numeric(RFACT),
    FSIZE = as.numeric(FSIZE),
    PUFEQ5 = as.numeric(PUFEQ5),
    PUFEQ6G = as.numeric(PUFEQ6G),
    TOINC = as.numeric(TOINC),
    PUFH04_SEX = as.numeric(PUFH04_SEX),
    PUFH05_AGE = as.numeric(PUFH05_AGE)
  )
```

#### All 5 survey years

```{r}
combined555 <- combined5 %>%
  filter(year %in% c(2011, 2013, 2014, 2016, 2017)) %>%
  mutate(year = factor(year))

combined555$year <- relevel(combined555$year, ref = "2016")
```

```{r}
event_model_all <- lm(
  FOOD ~ fourps * year +
         FSIZE +
         log(TOINC) +
         PUFEQ6G,
  data = combined555,
  weights = RFACT
)

summary(event_model_all) 
```

#### Graph

```{r}
#Extract coefficients:
library(broom)

event_results <- tidy(event_model_all)

#Filter interaction terms:
event_results <- event_results %>%
  filter(grepl("fourps:year", term))

#Create a year variable:
event_results$year <- c(2011, 2013, 2014,  2017)

#Plot the Event-Study
library(ggplot2)

# Plot the Event-Study at 10% significance level (90% CI)
ggplot(event_results, aes(x = year, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = estimate - 1.645*std.error, 
                    ymax = estimate + 1.645*std.error),
                width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    x = "Year",
    y = "Treatment Effect on Food Expenditure",
    title = "Event Study: Effect of Rice Subsidy on Food Expenditure"
  ) +
  theme_minimal()
```

```{r, echo=FALSE}
library(ggplot2)
library(scales)
library(broom)
library(dplyr)

# Extract and prepare results
event_results <- tidy(event_model_all) %>%
  filter(grepl("fourps:year", term)) %>%
  mutate(
    year = c(2011, 2013, 2014, 2017),
    # Label pre vs post
    period = ifelse(year == 2017, 
                    "Post-Intervention",
                    "Pre-Intervention"),
    # 90% CI bounds
    ci_low  = estimate - 1.645 * std.error,
    ci_high = estimate + 1.645 * std.error,
    # Significance flag
    significant = ifelse(ci_low > 0 | ci_high < 0,
                         "Significant", 
                         "Not Significant")
  )

# Plot
ggplot(event_results, 
       aes(x = year, y = estimate)) +
  
  # Pre/post shading
  annotate("rect",
           xmin = 2010.5, xmax = 2016.5,
           ymin = -Inf, ymax = Inf,
           fill = "gray95", alpha = 0.6) +
  annotate("rect",
           xmin = 2016.5, xmax = 2017.5,
           ymin = -Inf, ymax = Inf,
           fill = "lightyellow", alpha = 0.6) +
  
  # Intervention line
  geom_vline(xintercept = 2016.5,
             linetype = "dotted",
             color = "gray40",
             linewidth = 0.8) +
  
  # Zero reference line
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "gray30",
             linewidth = 0.8) +
  
  # Pre/post period labels
  annotate("text",
           x = 2013, y = max(event_results$ci_high) * 0.97,
           label = "Pre-Intervention",
           size = 3.2, color = "gray40",
           fontface = "italic") +
  annotate("text",
           x = 2017, y = max(event_results$ci_high) * 0.97,
           label = "Post",
           size = 3.2, color = "gray40",
           fontface = "italic") +
  
  # Reference year label
  annotate("text",
           x = 2016, 
           y = min(event_results$ci_low) * 0.85,
           label = "Reference\nYear (2016)",
           size = 2.8, color = "gray50",
           fontface = "italic") +
  
  # Reference year point
  annotate("point",
           x = 2016, y = 0,
           size = 3.5, shape = 21,
           fill = "white", color = "gray50",
           stroke = 1.2) +
  
  # Confidence intervals
  geom_errorbar(aes(ymin = ci_low,
                    ymax = ci_high,
                    color = significant),
                width = 0.3,
                linewidth = 0.9) +
  
  # Points colored by significance
  geom_point(aes(color = significant,
                 fill = significant),
             size = 4,
             shape = 21,
             stroke = 1.5) +
  
  # Colors
  scale_color_manual(
    values = c("Significant" = "#E07B6A",
               "Not Significant" = "#5B9BD5")
  ) +
  scale_fill_manual(
    values = c("Significant" = "#E07B6A",
               "Not Significant" = "#5B9BD5")
  ) +
  
  # Axis formatting
  scale_x_continuous(
    breaks = c(2011, 2013, 2014, 2016, 2017),
    labels = c("2011", "2013", "2014",
               "2016\n(Ref)", "2017\n(Post)")
  ) +
  scale_y_continuous(
    labels = comma_format(prefix = "₱")
  ) +
  
  # Labels
  labs(
    title = "Event Study: Effect of Rice Subsidy on Food Expenditure",
    subtitle = "Region X | Reference year: 2016 | 90% confidence intervals shown",
    x = "Survey Year",
    y = "Treatment Effect on Food Expenditure (₱)",
    color = "Statistical Significance",
    fill  = "Statistical Significance",
    caption = "Note: Pre-intervention estimates (2011, 2013, 2014) not significantly different from zero\nsupport the parallel trends assumption. Post-intervention estimate (2017) shows positive effect.\nSource: APIS 2011–2017, Philippine Statistics Authority."
  ) +
  
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold",
                              size = 12,
                              hjust = 0),
    plot.subtitle = element_text(size = 9,
                                 color = "gray40",
                                 hjust = 0),
    plot.caption = element_text(size = 8,
                                color = "gray50",
                                hjust = 0),
    axis.title = element_text(size = 11,
                              face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(face = "bold",
                                size = 10),
    legend.text = element_text(size = 10),
    legend.background = element_rect(
      fill = "white", color = "gray80"),
    panel.grid.major.y = element_line(color = "gray90",
                                      linewidth = 0.4),
    panel.grid.major.x = element_blank(),
    plot.margin = margin(15, 20, 15, 15)
  )

ggsave("event_study.png",
       width = 8, height = 5,
       dpi = 300, bg = "white")
```

## PART 6: DiD Estimation

```{r}
# Basic DiD
did_basic <- lm(FOOD ~ fourps * post,
                data = did_sample)
#summary(did_basic)

# With controls
did_controls <- lm(
  FOOD ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G,
  data = did_sample
)
#summary(did_controls)

# Weighted
did_weighted <- lm(
  FOOD ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G,
  data = did_sample,
  weights = RFACT
)
summary(did_weighted)

# Heteroskedasticity test
library(lmtest)
bptest(did_weighted)
#did_log <- lm(
#  log(FOOD) ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G,
#  data = did_sample,
#  weights = RFACT)
#library(lmtest)
#bptest(did_log)

# Robust SE
library(sandwich)
coeftest(did_weighted, vcov = vcovHC(did_weighted, type = "HC1"))
#library(sandwich)
#coeftest(did_log, vcov = vcovHC(did_log, type = "HC1"))

# ROBUST DiD MODEL for Reporting
#did_robust <- coeftest(did_weighted, vcov = vcovHC(did_weighted, type = "HC1"))

#did_data$resid_raw <- residuals(did_weighted)
#did_data$fitted_raw <- fitted(did_weighted)

#ggplot(did_data, aes(x = fitted_raw, y = resid_raw)) +
#  geom_point(alpha = 0.3) +
#  geom_hline(yintercept = 0, linetype = "dashed") +
#  labs(
#    title = "Residuals vs Fitted Values: Raw FOOD Model",
#    x = "Fitted Values",
#    y = "Residuals")
```

## PART 7: Robustness Checks

```{r}
# VIF
library(car)
vif(did_weighted)

# Winsorize FOOD
library(DescTools)
q_low  <- quantile(did_sample$FOOD, 0.01, na.rm = TRUE)
q_high <- quantile(did_sample$FOOD, 0.99, na.rm = TRUE)
did_sample$FOOD_w <- pmin(pmax(did_sample$FOOD, q_low), q_high)
did_winsor <- lm(
  FOOD_w ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G,
  data = did_sample,
  weights = RFACT)
summary(did_winsor)
coeftest(did_winsor, vcov = vcovHC(did_winsor, type = "HC1"))

# Log outcome
#did_log <- lm(
#  log(FOOD) ~ fourps * post + FSIZE + log(TOINC) + PUFEQ6G,
#  data = did_sample,
#  weights = RFACT)
#summary(did_log)

```

## SENSITIVITY ANALYSIS

### Robustness Value

```{r}
# ============================================================
# ROBUSTNESS VALUE ANALYSIS
# Rice Subsidy (2017) — APIS Region X
# Covariates: fourps:post, FSIZE, log(TOINC), PUFEQ6G
# ============================================================

library(sensemakr)
library(dplyr)
library(knitr)
library(kableExtra)


# ============================================================
# SECTION 1: sensemakr — Model 1 and 2
# ============================================================

# ── Model 1: Final weighted DiD (main model) ────────────────
sens_weighted <- sensemakr(
  model                = did_weighted,
  treatment            = "fourps:post",
  benchmark_covariates = c("FSIZE", "log(TOINC)", "PUFEQ6G"),
  kd    = 1,   # benchmark at 1× each covariate's R² with treatment
  ky    = 1,   # benchmark at 1× each covariate's R² with outcome
  q     = 1,   # test whether effect is reduced to zero
  alpha = 0.05,
  reduce = TRUE
)
ESV_weighted <- partial_r2(did_weighted, covariates = "fourps:post")
summary(sens_weighted)
summary(ESV_weighted)
plot(sens_weighted)
plot(sens_weighted, sensitivity.of = "t-value")



# ── Model 2: Basic DiD (no controls, no weights) ────────────
sens_basic <- sensemakr(
  model     = did_basic,
  treatment = "fourps:post",
  kd    = 1,
  ky    = 1,
  q     = 1,
  alpha = 0.05,
  reduce = TRUE
)
ESV_basic <- partial_r2(did_basic, covariates = "fourps:post")
summary(sens_basic)
summary(ESV_basic)
plot(sens_basic)
plot(sens_basic, sensitivity.of = "t-value")

```

```{r}
# ── Model 1 Plot Adjustments ──────────────────────────────────
# 1. Expand the X and Y plot limits slightly so edge labels don't get cut off
# 2. Use 'pos' or 'offset' rules if custom positioning is needed, or change label names
plot(sens_weighted, 
     lim = 0.4,             # Manually forces the R² bounds to expand/shrink for breathing room
     label.font = 2,        # Bold labels
     cex.lab = 0.8)         # Slightly shrink label text size to prevent overlap

plot(sens_weighted, 
     sensitivity.of = "t-value",
     lim = 0.4,
     cex.lab = 0.8)


# ── Model 2 Plot Adjustments ──────────────────────────────────
plot(sens_basic, 
     lim = 0.4,
     cex.lab = 0.8)

plot(sens_basic, 
     sensitivity.of = "t-value",
     lim = 0.4,
     cex.lab = 0.8)
```

```{r}
# ============================================================
# SECTION 3: Extract sensemakr RV for comparison
# ============================================================

extract_sensemakr_rv <- function(sens_obj,
                                 model_label,
                                 treatment_name = "fourps:post",
                                 bench_covariates = character(0)) {
  
  # Convert sensitivity stats safely
  stats <- as.data.frame(sens_obj$sensitivity_stats)
  
  # Main RV for treatment: point estimate
  rv_treat_point <- if ("rv_q" %in% names(stats)) {
    as.numeric(stats[["rv_q"]][1])
  } else {
    NA_real_
  }
  
  # Main RV for treatment: statistical significance
  rv_treat_sig <- if ("rva_q" %in% names(stats)) {
    as.numeric(stats[["rva_q"]][1])
  } else {
    NA_real_
  }
  
  rows <- data.frame(
    model             = model_label,
    covariate         = treatment_name,
    RV_point_estimate = round(rv_treat_point, 6),
    RV_stat_sig       = round(rv_treat_sig, 6)
  )
  
  # Benchmark RVs from bounds table
  if (length(bench_covariates) > 0 &&
      !is.null(sens_obj$bounds) &&
      nrow(as.data.frame(sens_obj$bounds)) > 0) {
    
    bounds <- as.data.frame(sens_obj$bounds)
    
    for (cov in bench_covariates) {
      
      row_name <- paste0("1x ", cov)
      
      matched_rows <- bounds[
        bounds$bound_label == row_name,
        ,
        drop = FALSE
      ]
      
      if (nrow(matched_rows) > 0) {
        
        rv_cov_point <- if ("rv_q" %in% names(matched_rows)) {
          as.numeric(matched_rows[["rv_q"]][1])
        } else {
          NA_real_
        }
        
        rv_cov_sig <- if ("rva_q" %in% names(matched_rows)) {
          as.numeric(matched_rows[["rva_q"]][1])
        } else {
          NA_real_
        }
        
        rows <- bind_rows(
          rows,
          data.frame(
            model             = model_label,
            covariate         = cov,
            RV_point_estimate = round(rv_cov_point, 6),
            RV_stat_sig       = round(rv_cov_sig, 6)
          )
        )
      }
    }
  }
  
  return(rows)
}

#bench_covs <- c("FSIZE", "log(TOINC)", "PUFEQ6G")

sens_rv_weighted <- extract_sensemakr_rv(
  sens_weighted,
  "Weighted DiD (Final)",
  treatment_name = "fourps:post",
#  bench_covariates = bench_covs
)

sens_rv_basic <- extract_sensemakr_rv(
  sens_basic,
  "Basic DiD",
  treatment_name = "fourps:post"
)


sensemakr_rv_all <- bind_rows(
  sens_rv_weighted,
  sens_rv_basic
)

sensemakr_rv_all


```
